# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data

```{r loaddata}
library(ggplot2)
activity <- read.csv("./activity.csv", header=TRUE, colClasses=c("integer", "Date", "integer"))
```


## What is mean total number of steps taken per day?

Splitting into daily data
```{r }
dailyActivity <- split(activity, factor(activity$date))
```

Totals, means and medians of the step count on daily basis
```{r }
totals <- sapply(dailyActivity, function(x) { sum(x$steps, na.rm=TRUE)})

dates <- names(totals)
names(totals) <- NULL

means <- sapply(dailyActivity, function(x) { mean(x$steps, na.rm=TRUE)})
names(means) <- NULL

medians <- sapply(dailyActivity, function(x) { median(x$steps, na.rm=TRUE)})
names(medians) <- NULL

dailyStats <- data.frame(dates, totals, means, medians)
print(dailyStats)

```
Histogram of daily total steps
```{r}
g <- ggplot(dailyStats , aes(dates, totals))
g + geom_bar(stat="identity")
```

## What is the average daily activity pattern?

Splitting into 5-minute-time-slot data
```{r }
slotActivity <- split(activity, factor(activity$interval))
```

Average step count on slot basis across all days
```{r }
averages <- sapply(slotActivity, function(x) { mean(x$steps, na.rm=TRUE)})
totals <- sapply(slotActivity, function(x) { sum(x$steps, na.rm=TRUE)})

slotStats <- data.frame(slots=as.numeric(names(averages)), averages=as.vector(averages), totals=as.vector(totals))
```

Time series plot

```{r}
qplot(slots, averages, data=slotStats, geom="line", group=1)
```

Slot containing the maximum steps

```{r}
maxValue <- max(slotStats$totals)
slotStats[slotStats$totals == maxValue, ]
```


## Imputing missing values

Report of total number of missing values
```{r}
dataWithNA <- activity[is.na(activity$steps), ]
nrow(dataWithNA)
```

Patching NA's with the mean
```{r}
for( i in 1:nrow(activity)) {
    if (is.na(activity[i, "steps"])) {
        slot <- activity[i, "interval"]
        averageValue <- slotStats[slotStats$slots == slot, "averages"]
        activity[i, "updated_steps"] <- averageValue
    } else {
        activity[i, "updated_steps"] <- activity[i, "steps"]
    }
}
```

Splitting into daily data (after patching)
```{r }
dailyActivity <- split(activity, factor(activity$date))
```

Totals, means and medians of the step count on daily basis
```{r }
totals <- sapply(dailyActivity, function(x) { sum(x$updated_steps, na.rm=TRUE)})

dates <- names(totals)

means <- sapply(dailyActivity, function(x) { mean(x$updated_steps, na.rm=TRUE)})

medians <- sapply(dailyActivity, function(x) { median(x$updated_steps, na.rm=TRUE)})

dailyStatsAfterPatch <- data.frame(dates=dates, totals=as.vector(totals), means=as.vector(means), medians=as.vector(medians))
```
Histogram of daily total steps after patch
```{r}
g <- ggplot(dailyStatsAfterPatch, aes(dates, totals))
g + geom_bar(stat="identity")
```
## Are there differences in activity patterns between weekdays and weekends? 

```{r}
activity$weekday <- weekdays(activity$date, abbreviate=TRUE)
for(i in 1:nrow(activity)) {
    day <- activity[i, "weekday"]
    if (day %in% c("Sat", "Sun")) {
        activity[i, "dayType"] <- "weekend"
    } else {
        activity[i, "dayType"] <- "weekday"
    }
}
```

Taking averages across weekdays and weekends
